---
output:
  word_document: default
  html_document: default
---
# Random Forests
### Augi, Alexandria

```{r, include=FALSE}
library(tidyverse)
library(tidymodels)
library(caret)
library(gridExtra)
library(vip)
library(ranger)
```

```{r}
drug = read_csv("drug_data-1.csv")
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
#str(drug)

drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"

drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44",
"45_54", "55_64", "65_"))) %>%
mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
mutate(Education = factor(Education, labels =
c("Under16", "At16", "At17", "At18", "SomeCollege",
"ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
mutate(Country = factor(Country,
labels = c("USA", "NewZealand", "Other", "Australia",
"Ireland","Canada","UK"))) %>%
mutate(Ethnicity = factor(Ethnicity,
labels = c("Black", "Asian", "White", "White/Black", "Other",
"White/Asian", "Black/Asian"))) %>%
mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
select(-ID)

#str(drug_clean)

drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
names(drug_clean)
```

### Task 1
```{r}
skimr::skim(drug_clean)
#str(drug_clean)
```

Using the skim package, there does not appear to be data missing from this set.

### Task 2
```{r}
set.seed(1234) 
drug_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine) 
train = training(drug_split)
test = testing(drug_split)

```

### Task 3
```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill")
p5 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4,p5)

p1 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
grid.arrange(p1,p2,p3, ncol = 2)

p1 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
grid.arrange(p1,p2,p3, ncol = 2)

ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()
```

It seems that the older ages, females, certain levels of education, ethnicity and the country all have impacts on the variable of Nicotine. Nscore and Oscore seem to have a slight effect bu the Escore does not. Cscore and impulsive have an impact on nicotine but the Ascore seems to be very slight. SS has a very noticeable impact on the nicotine variable.

### Task 4
```{r trouble}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)

Drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>%
  set_engine("ranger", importance = "permutation") %>%
  set_mode("classification")

Drug_wflow =
  workflow() %>%
  add_model(rf_model) %>%
  add_recipe(Drug_recipe)

set.seed(123)
rf_res = tune_grid(
  Drug_wflow,
  resamples = rf_folds,
  grid = 10)

rf_grid = grid_regular(
  mtry(range = c(2, 8)),
  min_n(range = c(5, 20)),
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  Drug_wflow,
  resamples = rf_folds,
  grid = rf_grid
)
rf_res %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")

rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")

```

### Task 5
```{r}
best_rf = select_best(rf_res_tuned, "accuracy")
final_rf = finalize_workflow(
  Drug_wflow,
  best_rf)

final_rf

final_rf_fit = fit(final_rf, train)

final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")

```

SS (Real) is sensation seeing measured by ImpSS seems to be the variable impacting nicotine the most. The country of residence, UK, follows the SS in importance level. Oscore seems reasonably important. The age range 45-54 of the respondent is important. Then the Nscore also appears to be important.

### Task 6
```{r}
trainpredrf = predict(final_rf_fit, train)
#head(trainpredrf)

confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")

testpredrf = predict(final_rf_fit, test)
#head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

On the training set, we have accuracy of 0.8945. Our no information rate is 0.6707. On the testing set, the accuracy is 0.7055 with the no information rate being 0.6702. These accuracy values are fairly lower than what we be preferred for the data set.

### Task 7
This model looking at the legal use of nicotine could be used for companies to determine what audiences to target advertisements to. They would be able to take this data and see the ages, gender, country, profession and other variables that determine the individuals use of nicotine and target products towards them. I do not believe I would recommend this model for "real world" use though. With the accuracy being below 90%, I think this model could be worked on more to improve that accuracy and then implemented in the real world.